Follow-up Samples

### 2017-02-20 submission/manuscript_outline_20161024.Rmd
results text corrections

If there were common OTUs that could separate adenoma and carcinoma from normal controls, we would expect to find a decrease in the positive probability of the follow up sample to be a lesion. This is what we observed for the lesion model (P-value `r ifelse(filter(all_probs_paired_wilcox, model_type =="red_lesion" & comparison == "lesion")[, "BH_correction"] < 0.001, paste("<"), paste(">"))` 0.001).  When we separated individuals based on whether they had an adenoma or carcinoma there was only a decrease in positive probability for the carcinoma group [Figure 3C] (P-value `r ifelse(filter(all_probs_paired_wilcox, model_type =="red_lesion" & comparison == "carcinoma_only")[, "BH_correction"] < 0.001, paste("<"), paste(">"))` 0.001) and not for the adenoma group [Figure 3D] (P-value `r ifelse(filter(all_probs_paired_wilcox, model_type =="red_lesion" & comparison == "all_adenoma")[, "BH_correction"] < 0.05, paste("<"), paste(">"))` 0.05).   We also observed that there were no significant differences in whether the models classified the samples as having lesion between the predicted and actual (P-value `r ifelse(filter(all_pred_stat_summary, samples_tested == "all" & model_type == "red_lesion" & info == "McnemarPValue")[, "all"] < 0.05, paste("<"), paste(">"))` 0.05).  The lesion model was able to correctly classify the one individual who still had a carcinoma on follow up [Figure 3C].

***The Initial Sample Model:***
After building a model to classify based on lesion we wanted to then build a separate model specifically to be able to classify whether samples were initial or follow up samples based on the bacterial community and FIT measurements.  The initial sample model had an AUC range of `r format(min(IF_AUC_run_summary$ROC), digits = 3)` to `r format(max(IF_AUC_run_summary$ROC), digits = 3)` after 100 iterations of 20 repeated 10-fold cross validations.  The test set AUC range for this model performed better then the training set AUCs.  There was a marked decrease in the ROC curve for the final model used when compared to the 100 test set AUC iterations [Figure 4A].  There were a total of `r length(IF_top10_varsMDA$variable)` variables that were used for this model [Figure 4B].  The variable that resulted in the largest MDA was *`r gsub("_unclassified", "", IF_imp_vars[as.character((filter(IF_top10_varsMDA, variable != "fit_result") %>% top_n(1, mean_MDA) %>% select(variable))[, 1]), "Genus"])`* (`r filter(IF_top10_varsMDA, variable != "fit_result") %>% top_n(1, mean_MDA) %>% select(variable)`) resulted in the largest decrease in MDA while FIT measurement for fecal blood resulted in the sixth largest decrease in MDA [Figure 4B].

>>>>

If there were common OTUs that could separate adenoma and carcinoma from normal controls, we would expect to find a decrease in the positive probability of the follow up sample to be a lesion. This is what we observed for the lesion model (P-value `r ifelse(filter(all_probs_paired_wilcox, model_type =="red_lesion" & comparison == "lesion")[, "BH_correction"] < 0.001, paste("<"), paste(">"))` 0.001).  When we separated individuals based on whether they had an adenoma or carcinoma there was only a decrease in positive probability for the carcinoma group [Figure 3C] (P-value `r ifelse(filter(all_probs_paired_wilcox, model_type =="red_lesion" & comparison == "carcinoma_only")[, "BH_correction"] < 0.001, paste("<"), paste(">"))` 0.001) and not for the adenoma group [Figure 3D] (P-value `r ifelse(filter(all_probs_paired_wilcox, model_type =="red_lesion" & comparison == "all_adenoma")[, "BH_correction"] < 0.05, paste("<"), paste(">"))` 0.05).   We also observed that there were no significant differences between the predicted and actual calls (P-value `r ifelse(filter(all_pred_stat_summary, samples_tested == "all" & model_type == "red_lesion" & info == "McnemarPValue")[, "all"] < 0.05, paste("<"), paste(">"))` 0.05).  The lesion model was also able to correctly classify the one individual who still had a carcinoma on follow up [Figure 3C].

***The Initial Sample Model:***
After building a model to classify based on lesion we wanted to then build a separate model specifically to be able to classify whether samples were initial samples based on the bacterial community and FIT measurements.  The initial sample model had an AUC range of `r format(min(IF_AUC_run_summary$ROC), digits = 3)` to `r format(max(IF_AUC_run_summary$ROC), digits = 3)` after 100 iterations of 20 repeated 10-fold cross validations.  The test set AUC range for this model performed better then the training set AUCs.  There was a marked decrease in the ROC curve for the final model used when compared to the 100 test set AUC iterations [Figure 4A].  There were a total of `r length(IF_top10_varsMDA$variable)` variables that were used for this model [Figure 4B].  The variable that resulted in the largest MDA was *`r gsub("_unclassified", "", IF_imp_vars[as.character((filter(IF_top10_varsMDA, variable != "fit_result") %>% top_n(1, mean_MDA) %>% select(variable))[, 1]), "Genus"])`* (`r filter(IF_top10_varsMDA, variable != "fit_result") %>% top_n(1, mean_MDA) %>% select(variable)`) while FIT measurement for fecal blood resulted in the sixth largest decrease in MDA [Figure 4B].

### 2017-02-20 submission/manuscript_outline_20161024.tex
results text corrections

If there were common OTUs that could separate adenoma and carcinoma from
normal controls, we would expect to find a decrease in the positive
probability of the follow up sample to be a lesion. This is what we
observed for the lesion model (P-value \textless{} 0.001). When we
separated individuals based on whether they had an adenoma or carcinoma
there was only a decrease in positive probability for the carcinoma
group {[}Figure 3C{]} (P-value \textless{} 0.001) and not for the
adenoma group {[}Figure 3D{]} (P-value \textgreater{} 0.05). We also
observed that there were no significant differences in whether the
models classified the samples as having lesion between the predicted and
actual (P-value \textgreater{} 0.05). The lesion model was able to
correctly classify the one individual who still had a carcinoma on
follow up {[}Figure 3C{]}.

\textbf{\emph{The Initial Sample Model:}} After building a model to
classify based on lesion we wanted to then build a separate model
specifically to be able to classify whether samples were initial or
follow up samples based on the bacterial community and FIT measurements.
The initial sample model had an AUC range of 0.641 to 0.805 after 100
iterations of 20 repeated 10-fold cross validations. The test set AUC
range for this model performed better then the training set AUCs. There
was a marked decrease in the ROC curve for the final model used when
compared to the 100 test set AUC iterations {[}Figure 4A{]}. There were
a total of 70 variables that were used for this model {[}Figure 4B{]}.
The variable that resulted in the largest MDA was \emph{Pseudomonas}
(Otu000438) resulted in the largest decrease in MDA while FIT
measurement for fecal blood resulted in the sixth largest decrease in
MDA {[}Figure 4B{]}.

>>>>

If there were common OTUs that could separate adenoma and carcinoma from
normal controls, we would expect to find a decrease in the positive
probability of the follow up sample to be a lesion. This is what we
observed for the lesion model (P-value \textless{} 0.001). When we
separated individuals based on whether they had an adenoma or carcinoma
there was only a decrease in positive probability for the carcinoma
group {[}Figure 3C{]} (P-value \textless{} 0.001) and not for the
adenoma group {[}Figure 3D{]} (P-value \textgreater{} 0.05). We also
observed that there were no significant differences between the
predicted and actual calls (P-value \textgreater{} 0.05). The lesion
model was also able to correctly classify the one individual who still
had a carcinoma on follow up {[}Figure 3C{]}.

\textbf{\emph{The Initial Sample Model:}} After building a model to
classify based on lesion we wanted to then build a separate model
specifically to be able to classify whether samples were initial samples
based on the bacterial community and FIT measurements. The initial
sample model had an AUC range of 0.641 to 0.805 after 100 iterations of
20 repeated 10-fold cross validations. The test set AUC range for this
model performed better then the training set AUCs. There was a marked
decrease in the ROC curve for the final model used when compared to the
100 test set AUC iterations {[}Figure 4A{]}. There were a total of 70
variables that were used for this model {[}Figure 4B{]}. The variable
that resulted in the largest MDA was \emph{Pseudomonas} (Otu000438)
while FIT measurement for fecal blood resulted in the sixth largest
decrease in MDA {[}Figure 4B{]}.
